---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# DataExtractCDCR

<!-- badges: start -->
<!-- badges: end -->

The goal of DataExtractCDCR is to enable the user to pull PDFs of data from the California Department of Corrections and Rehabilitation (CDCR) website, and convert that data to useable dataframes and files within R. 

## Installation

You can install the development version of DataExtractCDCR like so:

``` r
# FILL THIS IN! HOW CAN PEOPLE INSTALL YOUR DEV PACKAGE?
```

## Downloading PDFs


```{r}

# Pull the PDFs ========================================================================

source_pop <- "https://www.cdcr.ca.gov/research/2022-weekly-total-population-reports-tpop4/"
dest_pop <- "./data/raw/pop_reports/"

source_pop2 <- "https://www.cdcr.ca.gov/research/2021-weekly-total-population-reports-tpop4/"

source_compstat <- "https://www.cdcr.ca.gov/research/compstat-incident-reports-2021-2022/"
dest_compstat <- "./data/raw/compstat_reports/"

pull_pdfs(source_pop, dest_pop)
pull_pdfs(source_pop2, dest_pop)
pull_pdfs(source_compstat, dest_compstat)

```


## List the PDFs and Process 

```{r}

# list all of the PDF files
pdf_files_compstat <- list_pdfs("compstat_reports")
pdf_files_pop <- list_pdfs("pop_reports")

# there are some PDF files with an error, remove from processing for now
pdf_files_compstat <- pdf_files_compstat[-1]
#pdf_files_pop <- pdf_files_pop[-c(1, 2)]

compstat_dat <- bind_rows(lapply(pdf_files_compstat, pdf_to_df, type = "compstat"))  %>%
  # need to remove duplicate data that is contained in multiple PDFs
  distinct() %>%
  group_by(V1, month) %>%
  summarise(value = max(value))

pop_dat <- bind_rows(lapply(pdf_files_pop, pdf_to_df, type = "pop"))

saveRDS(compstat_dat, 
        paste0(here::here("data", "processed", "compstat_reports"), "/", Sys.Date(),".rds"))

saveRDS(pop_dat, 
        paste0(here::here("data", "processed", "pop_reports"), "/", Sys.Date(),".rds"))

```


You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/v1/examples>.


